---
title: "Calculate pre- and post-filtering cells"
output: html_notebook
---

This script counts cells for liver and PBMC objects from each woodchuck, after mitochondrial and DropletQC filtering (if applicable).

```{r}
library(Seurat)
library(scClustViz)
```

# Liver mito

Count cells in the liver that passed 50% mitochondrial content threshold. NOTE: this is after biopsy hepatocytes were removed.

Read in object

```{r}
sobj <- readRDS("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/no_dropletQC/allMerged_noDropletQC.RDS")
DimPlot(sobj, raster = FALSE)
```

```{r}
table(sobj$orig.ident)
```

# Liver DropletQC

Read in object

```{r}
sobj <- readRDS("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/dropletQC_filtered/allIntegrated_cca_kanchor5_noBiopsyHeps_dropletQCFiltered.RDS")
DimPlot(sobj)
```

```{r}
table(sobj$orig.ident)
```

# PBMC mito

Count PBMC cells from each woodchuck that passed the 10% mitochondrial content filter.

```{r}
load("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/no_dropletQC/integrated_PBMC_cca_kanchor5_scClustViz.RData")
DimPlot(scSeurat)
```

```{r}
table(scSeurat$orig.ident)
```

# Stimulated mito

```{r}
load("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/no_dropletQC/3391/integrated_3391_stimulation_cca_kanchor5_scClustViz.RData")
DimPlot(scSeurat)
```

```{r}
table(scSeurat$orig.ident)
```


