---
title: "R Notebook"
output: html_notebook
---

This script reads in both PBMC and Liver Seurat objects, and generates select figures used in the manuscript.

## Prepare objects

Load libraries

```{r}
library(Seurat)
library(scClustViz)
library(ggplot2)
library(dplyr)
```

Read in liver map

```{r}
sobj <- readRDS("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/dropletQC_filtered/allIntegrated_cca_kanchor5_noBiopsyHeps_dropletQCFiltered.RDS")
res <- "integrated_snn_res.1.4"
Idents(sobj) <- res
```

Read in PBMC map

```{r}
load("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/no_dropletQC/integrated_PBMC_cca_kanchor5_scClustViz.RData")
sobj <- scSeurat
res <- "integrated_snn_res.0.6"
Idents(sobj) <- res
```

## Visualization of metadata

UMAP with no cluster numbers

```{r}
DimPlot(sobj, label = FALSE) & NoLegend()
```

UMAP with cluster numbers

```{r}
DimPlot(sobj, label = TRUE)
```

Map with SCINA-generated cell-type labels

```{r}
DimPlot(sobj, group.by = "scina_labels_refined", label = TRUE) & NoLegend()
```

Map with original identities

```{r}
DimPlot(sobj, group.by = "orig.ident")
```

Barplot with original identities

```{r}
# Meta data to plot:
df <- sobj@meta.data
# Check what column the cluster identities are in
col <- which(colnames(df) == res)
# Order clusters
df[,col] <- factor(Idents(sobj),
                             levels = c(sort(as.numeric(levels(Idents(sobj))))))
# Basic plot of clusters by replicate
ggplot(df, aes(x = get(res), fill = orig.ident)) +
  geom_bar() +
  theme(axis.text = element_text(size = 7))
# Plot as proportion or percentage of cluster
ggplot(df, aes(x = get(res), fill = orig.ident)) +
  geom_bar(position = "fill") +
  theme(axis.text = element_text(size = 7))
```


## Dotplots

Generate dotplot with specific markers

```{r}
DotPlot(sobj,
        assay = "SCT",
        features = c("PTPRC", "CALCRL", "NKG7", "CD3E", "MARCO", "LYZ-1", "CD19", "MS4A1", "STAB2")
        ) +
  ggtitle("Select features for liver map")
```

## Feature plots

Make specific plots with specific genes. The genes we are interested in include: *PTPRC, CALCRL, NKG7, CD3E, MARCO, LYZ, CD19, MS4A1, STAB2, ALB, CD4, CD8A, CLEC4G, CD5L, C1QB, ACTA2, VWF, IGLL5, CD68*

```{r}
geneCode <- "sct_LYZ-1" # Woodchuck-specific nomenclature for this genome
gene <- "LYZ"
mapType <- "Liver"
FeaturePlot(sobj, features = geneCode) +
  ggtitle(paste(gene, "-", mapType, "map"))
```

Another version of the feature plot that outputs genes in italics

```{r}
FeaturePlot(sobj, features = geneCode) +
  ggtitle(expression(italic(gene)))
```

## Correlation plots

Generate heatmaps comparing woodchuck clusters with various datasets

### Setup for all correlations

First, set up whatever woodchuck dataset I am working with by reading in the ortholog table, choosing whether the orthologs to use are human, mouse, or woodchuck, and calculating the average expression for each cluster. The output of this section is the scaled cluster gene-expression matrix

```{r}
# Read in ortholog table
geneNameTable <- read.table("~/Dropbox/Zoe/scf_version/make_gtf/orthofinder_sc2/homologene/collectedOrthofinderPairings.tsv",
                            sep = "\t",
                            header = TRUE)
woodchuckClusterAverages <- AverageExpression(sobj,
                                              assays = "SCT",
                                              slot = "scale.data")
# Scale data
woodchuckClusterAverages$SCT <- na.omit(t(scale(t(woodchuckClusterAverages$SCT))))
# Grab gene names from Seurat object
uniqueHier <- row.names(woodchuckClusterAverages$SCT)
uniqueHier <- as.data.frame(uniqueHier)
# Bind with geneNameTable to get correct order (notice uniqueHier is on left)
newNames <- dplyr::left_join(uniqueHier, geneNameTable, by = "uniqueHier")
# Get orthologs from either mouse or human
species <- "human"
if (species == "human") {
  # If human one-to-one ortho has NA, replace with mikado_final_sc2_stringent_noMito_protein column
  # This is to avoid and potential mistakes in recognizing things it shouldn't be recognizing
  newNames$speciesOneToOne <- ifelse(is.na(newNames$humanOneToOne), newNames$uniqueHier, newNames$humanOneToOne)
} else if (species == "mouse") {
  newNames$speciesOneToOne <- ifelse(is.na(newNames$mouseOneToOne), newNames$uniqueHier, newNames$mouseOneToOne)
} else if (species == "woodchuck") {
  newNames$speciesOneToOne <- newNames$uniqueHier
}
# Grab dataframe
woodchuckClusterAverages <- woodchuckClusterAverages$SCT
# Replace names with one-to-one orthologue of particular species
row.names(woodchuckClusterAverages) <- newNames$speciesOneToOne
# Make sure formatted correctly
woodchuckClusterAverages <- as.data.frame(woodchuckClusterAverages)
# Order by gene name
woodchuckClusterAverages <- woodchuckClusterAverages[order(row.names(woodchuckClusterAverages)),]
```

### Dataset-specific setup

Correlation of woodchuck PBMCs with human 68k PBMC dataset from 10X Genomics

```{r}
# Now need to read in 68K PBMC data
humanPBMC <- read.csv("~/Dropbox/Zoe/scf_version/analysis/correlationTests/68K_pbmc_data/68K_enrichedGenes.csv",
                      header = FALSE)
# Get rid of top row
humanPBMC <- humanPBMC[-1,]
# Separate all cell and myloid cell data
allCells <- select(humanPBMC, V1, V2, V3)
myeloid <- select(humanPBMC, V5, V6, V7)
# Rename and get rid of first row
colnames(allCells) <- c("Cluster", "Gene", "Enrichment")
colnames(myeloid) <- c("Cluster", "Gene", "Enrichment")
# Get rid of top row
allCells <- allCells[-1,]
myeloid <- myeloid[-1,]
# I can then filter the rows and bind the data frames back together by gene name
for (j in 1:10) {
  dat <- dplyr::filter(allCells, Cluster == j)
  #dat <- get(paste("allCells", j, sep = ""))
  # Multiply enrichment values by -1 because the signs are backwards???
  dat$Enrichment <- as.numeric(dat$Enrichment) * -1
  dat <- dplyr::select(dat, Gene, Enrichment)
  colnames(dat) <- c("Gene", paste("Enrichment", j, sep = ""))
  if (j == 1) {
    allCellsMatrix <- dat
  }
  else {
    allCellsMatrix <- dplyr::full_join(allCellsMatrix, dat, by = "Gene")
  }
}
# Make row names gene names
rownames(allCellsMatrix) <- allCellsMatrix$Gene
allCellsMatrix <- dplyr::select(allCellsMatrix, -Gene)
# Make column names cell types
colnames(allCellsMatrix) <- c("Activated CD8+", "Naive CD8+", "Memory and Reg T",
                              "Naive CD4+", "NK", "CD8+", "B", "Megakaryocytes",
                              "Monocytes and Dendritic", "B, Dendritic, T")
# Scale across columns
allCellsMatrix <- t(scale(t(allCellsMatrix)))
# Order genes alphabetically by gene name
allCellsMatrix <- allCellsMatrix[order(row.names(allCellsMatrix)),]
```

Correlation of woodchuck liver with human liver dataset from MacParland *et al.* (2018)

```{r}
# Find cluster averages of human liver data
load("~/Dropbox/Zoe/scf_version/analysis/correlationTests/HumanLiver.RData")
# Run SCTransform
HumanLiverSeurat <- UpdateSeuratObject(HumanLiverSeurat)
HumanLiverSeurat <- SCTransform(HumanLiverSeurat)
humanClusterAverages <- AverageExpression(HumanLiverSeurat,
                                          assays = "SCT",
                                          slot = "scale.data")
# Replace cluster numbers with names
colnames(humanClusterAverages$SCT) <- c("Hep 1", "Alpha-beta T cells", "Hep 2",
                                        "Inflammatory macs", "Hep 3", "Hep 4",
                                        "Plasma cells", "NK-like cells", "Gamma-delta T cells",
                                        "Non-inflammatory macs", "Periportal LSECs", "Central venous LSECs",
                                        "Portal endothelial cells", "Hep 5", "Hep 6",
                                        "Mature B cells", "Cholangiocytes", "Gamma-delta T cells 2",
                                        "Erythroid cells", "Hepatic stellate cells")
# If only looking at specific clusters
#humanClusterAverages$SCT <- humanClusterAverages$SCT[,c("3","1","15","6","14","5")]
# Otherwise go straight to here:
humanClusterAverages$SCT <- na.omit(t(scale(t(humanClusterAverages$SCT))))
# Grab gene names
humanGenes <- row.names(humanClusterAverages$SCT)
# Now turn into large dataframe
allCellsMatrix <- as.data.frame(humanClusterAverages$SCT)
# Order by row name
allCellsMatrix <- allCellsMatrix[order(row.names(allCellsMatrix)),]
```

Correlation of woodchuck liver with human liver dataset from Aizarani *et al.*

```{r}
# Read in Aizarani dataset
aizarani <- readRDS("~/Dropbox/Zoe/scf_version/analysis/correlationTests/GSE124395_Normalhumanliverdata.RData")
# Read in clusters and label cells
aizaraniClusters <- read.table("~/Dropbox/Zoe/scf_version/analysis/correlationTests/GSE124395_clusterpartition.txt")
# Only keep cells in the cluster object
aizarani <- aizarani[,intersect(colnames(aizarani),row.names(aizaraniClusters))]
# Create Seurat object
aizarani <- CreateSeuratObject(counts = aizarani)
# Run SCTransform
aizarani <- SCTransform(aizarani)
# Add cluster IDs
Idents(aizarani) <- aizaraniClusters$sct.cpart
# Get cluster averages
aizaraniAverages <- AverageExpression(aizarani,
                                      assays = "SCT",
                                      slot = "scale.data")
aizaraniAverages$SCT <- na.omit(t(scale(t(aizaraniAverages$SCT))))
# Grab gene names
aizaraniGenes <- row.names(aizaraniAverages$SCT)
# Now turn into large dataframe
allCellsMatrix <- as.data.frame(aizaraniAverages$SCT)
# Order by row name
allCellsMatrix <- allCellsMatrix[order(row.names(allCellsMatrix)),]
```

Correlation of woodchuck liver with woodchuck PBMCs. For this correlation, read in the woodchuck liver dataset at the beginning of this script and then read in the woodchuck PBMCs below

```{r}
# Start with liver and read in woodchuck PBMCs again
load("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/no_dropletQC/integrated_PBMC_cca_kanchor5_scClustViz.RData")
Idents(scSeurat) <- "integrated_snn_res.0.6"
# Find cluster averages
pbmcClusterAverages <- AverageExpression(scSeurat,
                                         assays = "SCT",
                                         slot = "scale.data")
pbmcClusterAverages <- as.data.frame(na.omit(t(scale(t(pbmcClusterAverages$SCT)))))
# Order by row name
allCellsMatrix <- pbmcClusterAverages[order(row.names(pbmcClusterAverages)),]
```

### Output plots for all correlations

```{r}
speciesData <- "Human Liver (Aizarani et al.)"
woodchuckData <- "Woodchuck Liver"
# Now find intersecting genes
matches <- intersect(row.names(allCellsMatrix),
                     row.names(woodchuckClusterAverages))
# Look at how many genes matched
length(matches)
# Make new matrices with only matching gene names
toCor <- allCellsMatrix[matches,]
woodchuckAveragesCor <- woodchuckClusterAverages[matches,]
# Do Pearson
pearVal <- cor(toCor, woodchuckAveragesCor, method = "pearson")
heatmap(pearVal,
        main = paste("Pearson correlation of", speciesData, "vs", woodchuckData),
        xlab = woodchuckData,
        ylab = speciesData)
        #margins = c(6,11))
#Rowv = NA,
#Colv = NA)
# Do Spearman
spearVal <- cor(toCor, woodchuckAveragesCor, method = "spearman")
heatmap(spearVal,
        main = paste("Spearman correlation of", speciesData, "vs", woodchuckData),
        xlab = woodchuckData,
        ylab = speciesData)
        #margins = c(6,11))
#Rowv = NA,
#Colv = NA)
```




