---
title: "R Notebook"
output: html_notebook
---

Analysis and visualisation of spatial transcriptomics data. Generally followed [this Seurat workflow](https://satijalab.org/seurat/archive/v3.2/spatial_vignette.html).

Load libraries

```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
library(UCell)
#library(ggpubr)
```

Load merged liver slices and set identity

```{r}
sobj <- readRDS("../spatial/merged_spatial_sobj.RDS")
Idents(sobj) <- "SCT_snn_res.0.2"
```

Rename clusters

```{r}
clustNames <- Idents(sobj)
clustNames <- gsub(0, "Cluster 0 (Periportal)", clustNames)
clustNames <- gsub(1, "Cluster 1 (Pericentral)", clustNames)
clustNames <- gsub(2, "Cluster 2 (Immune)", clustNames)
clustNames <- gsub(3, "Cluster 3 (Immune)", clustNames)
clustNames <- gsub(4, "Cluster 4 (Immune)", clustNames)
sobj$general_cell_labels <- clustNames
SpatialDimPlot(sobj, group.by = "general_cell_labels")
```


Visualise clusters on spatial plot

```{r}
plot1 <- SpatialDimPlot(sobj, group.by = "general_cell_labels", alpha = 0.7, images = 'slice1') & NoLegend()
plot2 <- SpatialDimPlot(sobj, group.by = "general_cell_labels", alpha = 0.7, images = 'slice1.1') +
  labs(fill = "Cluster")
plot1 + plot2
pdf("./figures/spatial/merged_dimPlot_UMAP.pdf")
plot1 + plot2
dev.off()
```

Visualise features on spatial plot

```{r}
genes <- c("CYP2E1","FETUB","HMGCS1","APOC2","Saa2;Saa1-1","HAMP")
for (gene in genes) {
  plot1 <- SpatialFeaturePlot(sobj, features = gene, images = 'slice1') & NoLegend()
  plot2 <- SpatialFeaturePlot(sobj, features = gene, images = 'slice1.1') & NoLegend()
  print(plot1 + plot2)
  pdf(paste("./figures/spatial/merged_spatialPlot_", gene, ".pdf", sep = ""))
  print(plot1 + plot2)
  dev.off()
}
```

Look at other genes

```{r}
gene <- "Saa2;Saa1-1"
gene <- "RSPO3"
gene <- "MECOM"
gene <- "KRT19"
gene <- "EPCAM"
gene <- "SOX9"
gene <- "ANXA2"
gene <- "BIRC3"
plot1 <- SpatialFeaturePlot(sobj, features = gene, images = 'slice1') & NoLegend()
plot2 <- SpatialFeaturePlot(sobj, features = gene, images = 'slice1.1') & NoLegend()
print(plot1 + plot2)
```


Calculate markers

```{r}
all.markers <- FindAllMarkers(sobj, 
                              only.pos = TRUE, 
                              min.pct = 0.25, 
                              logfc.threshold = 0.25,
                              recorrect_umi = FALSE)
top.markers <- all.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
top.markers
write.table(all.markers,
            file = "./figures/spatial/merged_spatial_markers.tsv",
            sep = "\t",
            quote = FALSE,
            row.names = FALSE)
```

Create zonation scores using top 10 genes from periportal and pericentral clusters

```{r}
# Compare clusters 0 and 1
markers <- FindMarkers(sobj,
                       ident.1 = '0',
                       ident.2 = '1',
                       only.pos = FALSE,
                       min.pct = 0.25,
                       logfc.threshold = 0.25,
                       recorrect_umi = FALSE)
# Remove mikado genes
markers <- markers[grep("mikado", rownames(markers), invert = TRUE),]
# Get top 30 periportal genes
pp_genes <- rownames(markers[markers$avg_log2FC > 0,])[1:10]
pp_genes
# Get top 30 pericentral genes
cv_genes <- rownames(markers[markers$avg_log2FC < 0,])[1:10]
cv_genes
```

Format for UCell

```{r}
mySigs <- list()
mySigs$Periportal <- pp_genes
mySigs$Pericentral <- cv_genes
mySigs
```


Add these as gene signatures

```{r}
sobj <- AddModuleScore_UCell(sobj, features = mySigs, assay = "SCT")
signature.names <- paste0(names(mySigs), "_UCell")
VlnPlot(sobj, features = signature.names, group.by = "SCT_snn_res.0.2")
SpatialFeaturePlot(sobj,
                   features = signature.names,
                   images = 'slice1')
SpatialFeaturePlot(sobj,
                   features = signature.names,
                   images = 'slice1.1')
```

Create figures of spatial plots with signatures

```{r}
plot1 <- SpatialFeaturePlot(sobj,
                            features = "Periportal_UCell",
                            images = 'slice1')
plot2 <- SpatialFeaturePlot(sobj,
                            features = "Periportal_UCell",
                            images = 'slice1.1')
pdf("./figures/spatial/periportalSig_spatial_UCell.pdf")
plot1 + plot2
dev.off()
```

```{r}
plot1 <- SpatialFeaturePlot(sobj,
                            features = "Pericentral_UCell",
                            images = 'slice1')
plot2 <- SpatialFeaturePlot(sobj,
                            features = "Pericentral_UCell",
                            images = 'slice1.1')
pdf("./figures/spatial/pericentralSig_spatial_UCell.pdf")
plot1 + plot2
dev.off()
```

Create heatmap with signature genes

```{r}
plot <- DoHeatmap(sobj,
                  features = c(pp_genes, cv_genes),
                  label = FALSE,
                  size = 3,
                  group.by = "general_cell_labels")
plot
pdf("./figures/spatial/zoneSigs_byClust_heatmap.pdf")
plot
dev.off()
```

Correlate UMAP zonation scores with gene expression

```{r}
FeaturePlot(sobj, features = c("WNT2", "Pericentral_UCell"), blend = TRUE)
FeatureScatter(sobj, feature1 = "WNT2", feature2 = "Pericentral_UCell")
```

Calculate Z-score for each gene for CV and PP signatures

```{r}
zscore = (sobj@assays$SCT@scale.data['CYP2E1',] - sobj$Pericentral_UCell)/sd(sobj$Pericentral_UCell)
sobj$cyp2e1_zscore <- scale(zscore)
SpatialFeaturePlot(sobj, features = 'cyp2e1_zscore', images = 'slice1')
```

Try with RSPO3

```{r}
zscore = (sobj@assays$SCT@scale.data['RSPO3',] - sobj$Pericentral_UCell)/sd(sobj$Pericentral_UCell)
sobj$rspo3_zscore <- scale(zscore)
SpatialFeaturePlot(sobj, features = 'rspo3_zscore', images = 'slice1')
```

Correlate all genes with CV and PP signatures

```{r}
cv_cor <- apply(sobj@assays$SCT@counts, 1, function (x) cor(x, y = sobj$Pericentral_UCell,
                                                                method = "spearman"))
head(sort(cv_cor, decreasing = TRUE))
pp_cor <- apply(sobj@assays$SCT@counts, 1, function (x) cor(x, y = sobj$Periportal_UCell,
                                                                method = "spearman"))
head(sort(pp_cor, decreasing = TRUE))
```

Write table

```{r}
spatial_cors <- cbind(cv_cor, pp_cor)
write.table(spatial_cors,
            file = "./figures/spatial/spatial_gene_correlations.tsv",
            quote = FALSE,
            sep = "\t",
            row.names = TRUE,
            col.names = TRUE)
```


Save object if needed

```{r}
saveRDS(sobj,
        file = "../spatial/merged_spatial_sobj.RDS")
```


