---
title: "Stimulation experiments"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

Analysis of stimulation experiments.

```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
```

Load integrated object. This object has been integrated with CCA, because there was lots of batch effect between the 0hr tissue and the other two experiments (which merged quite well).

```{r}
load("~/Dropbox/Zoe/scf_version/analysis/healthy_sc/seurat_objects/no_dropletQC/3391/integrated_3391_stimulation_pmaio_unstim_downsampled_cca_kanchor5_scClustViz.RData")
```

Set identities:

```{r}
Idents(scSeurat) <- "integrated_snn_res.0.4"
```


Check out the UMAP of clusters, original identities, and refined SCINA labels.

```{r}
DimPlot(scSeurat, label = TRUE)
DimPlot(scSeurat, group.by = "tissue_type", label = FALSE)
DimPlot(scSeurat, group.by = "scina_labels", label = TRUE) & NoLegend()
```

Look at tissue-type per cluster on a proportional level:

```{r}
pt <- table(Idents(scSeurat), scSeurat$tissue_type)
pt <- as.data.frame(pt)
ggplot(pt, aes(x = Var1, y = Freq, fill = Var2)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  theme(legend.title = element_blank())
```

Look for T-cell signatures.

```{r}
FeaturePlot(scSeurat, features = c("CD3E", "NKG7"))
FeaturePlot(scSeurat, features = c("LEF1", "XCL1;XCL2"))
FeaturePlot(scSeurat, features = c("TOX", "CD8A"))
```

Look for macrophage signatures:

```{r}
FeaturePlot(scSeurat, features = c("LGALS1", "TYROBP"))
FeaturePlot(scSeurat, features = c("MARCO", "C1QC"))
FeaturePlot(scSeurat, features = c("LYZ-1", "S100A8"))
FeaturePlot(scSeurat, features = c("DEFA6;DEFA4;DEFA5", "CTSS"))
FeaturePlot(scSeurat, features = c("FLT3", "S100A12"))
```

Look for B and plasma cell signatures

```{r}
FeaturePlot(scSeurat, features = c("CD79B", "JCHAIN"))
```


Cluster 12 is the only cluster that is obviously dominated by the unstimulated experiments. What cells are these? Looks like they are hepatocytes. A lot of the markers are periportal, but all hepatocytes are in this cluster.

```{r}
# Isolate DE genes
clust11 <- sCVdata_list$res.0.4@DEvsRest$`12`[order(sCVdata_list$res.0.4@DEvsRest$`12`$Wstat,
                                                    decreasing = TRUE), ]
rownames(clust11)[1:50]
FeaturePlot(scSeurat, features = c("PAH","PCK1"))
FeaturePlot(scSeurat, features = c("CYP2E1","FETUB"))
FeaturePlot(scSeurat, features = c("ACACB","ELOVL6"))
FeaturePlot(scSeurat, features = c("POLR2D", "ALB-1"))
```

Let's look at some of the key markers that Sonya sent: IL2, IL4, IL6, IL7, IL10, IL12A, TGFB

```{r}
FeaturePlot(scSeurat, features = c("IL2", "IL4"))
FeaturePlot(scSeurat, features = c("IL6", "IL6-1"))
FeaturePlot(scSeurat, features = c("IL7", "IL10"))
FeaturePlot(scSeurat, features = c("IL12A", "TGFB1"))
```

It's pretty hard to see which cytokines are expressed in which samples, so I'm going to try splitting up the samples.

```{r}
DimPlot(scSeurat, split.by = "tissue_type", label = TRUE) & NoLegend()
```

Now let's take a look at the features on this split plot.

```{r}
FeaturePlot(scSeurat, feature = "IL2", split.by = "tissue_type") # Only in PMAIO cluster 2
FeaturePlot(scSeurat, feature = "IL4", split.by = "tissue_type") # Only in PMAIO cluster 2
FeaturePlot(scSeurat, feature = "IL6", split.by = "tissue_type") # Stronger in endo and mesem PMAIO
FeaturePlot(scSeurat, feature = "IL6-1", split.by = "tissue_type") # No signal
FeaturePlot(scSeurat, feature = "IL7", split.by = "tissue_type") # More in PMAIO cluster 4/5
FeaturePlot(scSeurat, feature = "IL10", split.by = "tissue_type") # More in PMAIO cluster 2
FeaturePlot(scSeurat, feature = "IL12A", split.by = "tissue_type") # No signal
FeaturePlot(scSeurat, feature = "TGFB1", split.by = "tissue_type") # Pretty even across all
FeaturePlot(scSeurat, feature = "IFNG", split.by = "tissue_type") # More in PMAIO cluster 2
FeaturePlot(scSeurat, feature = "TGFA", split.by = "tissue_type")
FeaturePlot(scSeurat, feature = "TNF", split.by = "tissue_type")
```

Cluster 2 looks like it generally has more IL genes expressed. Isolate DE genes for this cluster.

```{r}
clust1 <- sCVdata_list$res.0.4@DEvsRest$`2`[order(sCVdata_list$res.0.4@DEvsRest$`2`$Wstat,
                                                    decreasing = TRUE), ]
cat(rownames(clust1)[1:50])
```

Visualise these genes in violin plots.

**Interpretation**: Now I see that IL2 and IL4 are expressed in more than just cluster 2, but it is nearly always in PMAIO. Cluster 8 is definitely where IL6 is expressed. IL7 is a bit more spread out across treatment types, but it's definitely present in PMAIO in clusters 4, 5, and 11. Cluster 11 is plasma cells, but I'm honestly not sure of 4 and 5 - low signal hepatocytes? IL10 is in clusters 2 and 11 PMAIO. IL12A is a bit more in 4 and 5 PMAIO.

```{r}
VlnPlot(scSeurat, feature = "IL2", split.by = "tissue_type") # Only in PMAIO cluster 2
VlnPlot(scSeurat, feature = "IL4", split.by = "tissue_type") # Only in PMAIO cluster 2
VlnPlot(scSeurat, feature = "IL6", split.by = "tissue_type") # Stronger in endo and mesem PMAIO
VlnPlot(scSeurat, feature = "IL6-1", split.by = "tissue_type") # No signal
VlnPlot(scSeurat, feature = "IL7", split.by = "tissue_type") # More in PMAIO cluster 4/5
VlnPlot(scSeurat, feature = "IL10", split.by = "tissue_type") # More in PMAIO cluster 2
VlnPlot(scSeurat, feature = "IL12A", split.by = "tissue_type") # A bit more in 1,4,5
VlnPlot(scSeurat, feature = "TGFB1", split.by = "tissue_type") # Pretty even across all
VlnPlot(scSeurat, feature = "TNF", split.by = "tissue_type")
```


